---
title: "Report of the Severe Weather impact in US's health and economy"
author: "zhenyue zhu"
date: "May 8, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

Storms and other severe weather events can cause both public health and economic 
problems for the world. They can result in fatalities, injuries, and property 
damage from 1950 to present. Thus it is crucial for us to understand what kind of 
events have the strongest effect for our health and economy.

This project involves exploring the U.S. National Oceanic and Atmospheric 
Administration's (NOAA) storm database. This database tracks characteristics of 
major storms and weather events in the United States, including when and where they 
occur, as well as estimates of any fatalities, injuries, and property damage.

The purpose of this project is address the following questions:

* Across the United States, which types of events are most harmful with respect to
population health?
* Across the United States, which types of events have the greatest economic 
consequences?

## Data Processing
First download the data from the website and then directly read the bz2 format file 
into the data set.

```{r cars, cache=TRUE}
library(ggplot2)
url="http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(url, destfile = "~/data/R/stormData.csv.bz2")
data<-read.csv("~/data/R/stormData.csv.bz2")
```

Get some basic information of this data set.
```{r, cache=TRUE}
dim(data)
#head(data,n=10)
str(data)
summary(data$EVTYPE)
class(data$EVTYPE)
```

As we can see that this event types has many things which have very similar names.
For example, "THUNDERSTORM WIND" and "THUNDERSTORM WINDS" should be similar things.
Then I will define the function which will make similar event types to be similar.

```{r,cache=TRUE}
reduce_cateory<-function(x){
  x <- tolower(x)
  if (grepl("hail", x)){
    return("hail")
  }
    if (grepl("flood|FLD|fld", x)){
    return("flood")
  }
  if (grepl("thunderstorm", x)){
    return("thunderstorm")
  }
  if (grepl("wind", x)){
    return("wind")
  }
  if (grepl("dry|drought", x)){
    return("dry weather")
  }
  if (grepl("heat|warm", x)){
    return("hot weather")
  } 
  if (grepl("fire", x)){
    return("Fire")
  }
  if (grepl("rain", x)){
    return("rain")
  }
  if (grepl("hurricane|tropical|typhoon", x)){
    return("hurricane")
  }
  if (grepl("snow|ice|freez|blizzard|winter", x)){
    return("winter disaster")
  }
  else{
    return(x)
  }
}

events<-as.factor(sapply(data$EVTYPE,FUN=reduce_cateory))
summary(events)
```
Now We see that these event types are greatly reduced. 
The most common types are "hail","wind","thunderstorm","flood","tornado","winter 
disaster" and "lightning". I will create the damage to the health and the damage to 
the crops.

## consider the events which has the most injuries or fatalities

Then I will sum the fatalties and injuries for these type events and create the total data frame.
```{r,cache=TRUE}
fatal <- aggregate(data$FATALITIES, list(events), sum)
injury <- aggregate(data$INJURIES, list(events), sum)
names(fatal) <- c("event", "count"); 
names(injury) <- c("event", "count")
health <- data.frame(fatal$event, fatal$count, injury$count)
names(health) <- c("event", "fatalities", "injuries")
health <- health[order(-health$injuries, -health$fatalities), ][1:10,]
health
```
It's very obvious from this table that the max fatalities type from 1950 to present 
is tornado related disasters. Similarly, 
the most injuries types are also related to tornados.
```{r}
library(ggplot2)
qplot(data = health, x = event, y = fatalities, geom = c("point", "smooth"))+
labs(title = '10 Most Harmful Types of Weather\n Measured by Fatalities from 1950 - 
     present', y = 'Number of People', x = 'Weather Event')
```

## Consider the event which has the most influence on the economic consequences.
I need the columns with "PRODMG" and "CROPDMG", since these columns have different units, I have to convert the units from k to 10^3, M to 10^6, etc...

```{r,cache=TRUE}
convert <- function(coeff, exp){
  if (is.na(exp)){
    as.numeric(coeff)
  }
  else if (toupper(exp)== "K"){
    as.numeric(coeff)*10^3
  }
  else if (toupper(exp) == "M"){
    as.numeric(coeff)*10^6
  }
  else if (toupper(exp)== "B"){
    as.numeric(coeff)*10^9
  }
  else{
    as.numeric(coeff)
  }
}

prop <- apply(data[, c('PROPDMG', 'PROPDMGEXP')], 1, function(x) convert(x['PROPDMG'], x['PROPDMGEXP']))
crop <- apply(data[, c('CROPDMG', 'CROPDMGEXP')], 1, function(x) convert(x['CROPDMG'], x['CROPDMGEXP']))
```

```{r,cache=TRUE}
prop_dmg <- aggregate(prop, list(events), sum)
crop_dmg <- aggregate(crop, list(events), sum)
names(prop_dmg) <- c("event", "cost"); 
names(crop_dmg) <- c("event", "cost")
economy <- data.frame(prop_dmg$event, prop_dmg$cost, crop_dmg$cost)
names(economy) <- c("event", "prop", "crop")
economy <- economy[order(-economy$prop, -economy$crop), ][1:10,]
economy
```

```{r}
library(ggplot2)
qplot(data = economy, x = event, y = prop/10^9, geom = c("point", "smooth"))+
labs(title = '10 Most Harmful Types of Weather\n Measured by property damage from 
     1950 - present', y = 'in billions of US dollars', x = 'Weather Event')
```

It's easy to see that flood has the highest both property and crop damage.

## Results and Conclusion.

* The most dangerous event type for the people's health is **Tornado**.
* The most property and crop damage event type is **flood**. 
