---
title: "ToothGrowth Analysis"
author: "zhenyue zhu"
date: "April 24, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The purpose of this project is to analyze the ToothGrowth data in the R datasets 
package. The data includes the length of odontoblasts (cells responsible for tooth 
growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin 
C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC). 

##Load the ToothGrowth data and perform some basic exploratory data analyses.

```{r}
library(ggplot2)
data("ToothGrowth")
str(ToothGrowth)
summary(ToothGrowth)
par(mfrow = c(1,2))
p1 <- ggplot(ToothGrowth, aes(x = factor(dose), y = len, fill = factor(dose))) 
p1 + geom_boxplot() + guides(fill=FALSE) + facet_grid(. ~ supp)
p2 <- ggplot(ToothGrowth, aes(x = factor(supp), y = len, fill = factor(supp))) 
p2 + geom_boxplot() + guides(fill=FALSE) + facet_grid(. ~ dose)
```

After plot the data, we see that for dose versus length, there is always an 
increasing trend, i.e. more dose, the tooth length will increase faster.
But it looks like when a small dose is applied, the dose of OJ(orange juice) will 
make the teeth grow faster. At dose 2.0, it looks like both supplements has the 
similar effect on the tooth growth.

## Statistical inferences
The tooth growth was compared by supplement at different dose, under the null 
hypothesis that each supplement has the same effect at the same dose for tooth 
growth. Then we divided the data sets into three doses groups. Then perform a 
t-test comparing length and supplements within each group. 

```{r}
df0.5 <- subset(ToothGrowth, dose == 0.5)
df1.0 <- subset(ToothGrowth, dose == 1.0)
df2.0 <- subset(ToothGrowth, dose == 2.0)

test0.5 <- t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = df0.5)
test0.5$p.value 
test0.5$conf

test1.0 <- t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = df1.0)
test1.0$p.value 
test1.0$conf
```
For group dose is 0.5 or 1.0, we see that the p-value is very small (0.0064 and 
0.0010 for dose 0.5 and 1.0). Confidence intervals are both positive. 
Indicating that the mean between each supplement is significant. 
```{r}
test2.0 <- t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = df2.0)
test2.0$p.value 
test2.0$conf
```
However, for group with dose=2.0, the p-value is very high 0.96 and confidence 
interval includes negative. It indicate that for dose=2.0 group, different supplement has negligible effect on the tooth growth.

## Assumptions.
I based the above inferences on the facts that:

* only orange juice and VC has effect on the tooth growth, without any other 
factors. 
* 60 different samples are unrelated amd different samples has different variance. That's why I use paired=False and var.equal=False.
* The suject has identical other factors. 

## Conclusion.
Different supplements (VC and OJ) has different effects on tooth growth at lower 
dose according to t-test. At dose=0.5 and 1.0, OJ has stronger effect on the teeth 
growth. While at dose=2.0, there is no significant difference in tooth length 
comparing two supplements. 